Online Balanced Graph Partitioning: Optimizations for Large Clusters

### 2020-05-01 paper/disc20clustering.tex
Non-formal definitions in the lower bound

We say that a component or a ground set is \emph{trivial}, if it contains a single node.
We say that a~configuration is \emph{component respecting}
if the nodes belonging to the same component are collocated.
Similarly, we define a \emph{ground-set-respecting} configuration.

@@

\begin{theorem}
	\label{th:lowerbound}
	The competitive ratio of any deterministic online algorithm for \OBRP{} is  $(k \cdot \ell - k - \ell) / 2 = \Omega(k\cdot \ell)$ for any $k\geq 3$.
\end{theorem}

\begin{proof}
	We construct an instance of the problem with $\ell$ clusters 
	$\set{ S_1, S_2,\dots , S_{\ell}}, |S_i|  = k$.
	Let $I(C)$ denote the cluster where nodes of component $C$ are located initially.
	%	\maciek{So far components were not introduced; the best place to introduce these would be around the definition of \PPOBRP}.
	Fix any online algorithm \ALG{}.

	\maciek{This paragraph should be before the proof, merged with section about ground sets (i.e., adversary components.)}
	We join some nodes into components, and produce the input sequence by observing actions of \ALG{} with respect to nodes of these components.
	If \ALG{} at any point splits a component
	(i.e., spreads its nodes over two or more clusters),
	then we continue to issue requests between non-collocated nodes of the component until \ALG{} collocates the nodes of the component.
	If \ALG{} never collocates them, then it is not competitive.

	Initially each node belongs to its own component of size $1$.
	We say that a node is single if it belongs to a singleton component.
	We begin by forming a component of single $k-1$ nodes on the cluster $S_1$.

	Let $x_0$  be the only single node left on $S_1$ and  $y_0 \in S_2$ be any single node on $S_2$.
	Next,
	we merge components of $x_0$~and~$y_0$.
	Since these reside on different clusters of \ALG{}, it produces an external request.
	\ALG{} must collocate them in a~cluster other than $S_1$.
	To this end,
	\ALG{} moves to a new configuration
	that replaces $x_0$ and $y_0$ with two other single nodes $x_1$ and $y_1$ respectively.

>>>>

We say that a component or a ground set is \emph{trivial}, if it contains exactly one node.
We say that the node that belongs to trivial component or a ground set is single.
We say that a~configuration is \emph{component respecting}
if the nodes belonging to the same component are collocated.
Similarly, we define a \emph{ground-set-respecting} configuration.

@@

\begin{theorem}
	\label{th:lowerbound}
	The competitive ratio of any deterministic online algorithm for \OBRP{} is  $(k \cdot \ell - k - \ell) / 2 = \Omega(k\cdot \ell)$ for any $k\geq 3$ and for any $\ell \geq 2$.
\end{theorem}

\begin{proof}
	Fix any online algorithm \ALG{}.
	Let $I(C)$ denote the cluster where nodes of ground set $C$ are located initially.
	\maciek{The problem with the definition! The first pair does not have the same initial location of its nodes.}
	Initially all nodes are single (they belong to their own trivial ground sets).

	In our construction we distinguish a cluster $S_1$, where we form a ground set of size $k-1$.
	Let $x_0$  be the only single node left on $S_1$ and let $y_0$ be an  arbitrary node that does not reside in $S_1$.
	First,
	we construct a ground set $\{x_0, y_0\}$.
	Since \ALG{} has them split, the adversary produces an external request.
	Since $S_1$ contains a ground set of size $k$, \ALG{} must collocate $\{x_0, y_0\}$ in another cluster.
	To preserve a feasible partition of nodes into clusters,
	 replaces $x_0$ and $y_0$ with two other single nodes $x_1$ and $y_1$ respectively.
	 \maciek{Not covering all cases. $y_0$ may stay in its place. This sentence should be about $x_0$ replaced by $x_1$ only.}
